Correspondence in Significance

### 2018-05-21 significance/significance.Rmd
Updated significance writeup

Programs of research that examine the replicability of research findings have become more common in psychology. 
However, there is some ambiguity over how to assess the findings of such programs; that is, how do we know if a finding has been replicated?
The Open Science Collaboration (2015) argue, "There is no single standard for evaluating replication success," and then analyzed replicate pairs in a variety of ways.
Subsequent programs have used some of these same analysis methods (e.g., Camerer, et al., 2016; Schweinsberg, et al., 2016).

It is not entirely clear what these methods say about replication.
For one, an actual definition of replication remains somewhat elusive---or flexible---across these methods. 
Saying that studies get "the same result" is somewhat subjective, and the OSC attempted to operationalize this by describing several different scenarios of studies getting "the same result".
Moreover, interpreting the results of the OSC analyses also requires knowledge of the properties of the methods they (and others) use.
For instance, how sensitive are these analyses? 
The irony is that answering this second question, in some ways, requires an answer to the former: how do we define replication?

In this paper, we clarify subjective notions about replication using a meta-analytic framework. 
Given this definition, we then evaluate various analysis methods to determine their properties.

# Defining Successful Replication: Meta-analytic Perspectives

Meta-analysis represents the "results" of an experiment by an effect parameter, denoted $\theta$.
This "true" effect reflects what the researcher would observe if they had an infinite sample size.
Effect parameters are typically on a standard scale, such as a standardized mean difference, Fisher-transformed correlation coefficient, or log-odds ratio.

In the context of replication studies, we have at least $k \geq 2$ such findings, so let $\theta_i$ describe the results of the $i$th study.
For the RPP and RPE programs, this invovles $\theta_1$ for the original study and $\theta_2$ for the replicate.
However, we do not actually observe $\theta_1$ and $\theta_2$, and instead we estimate them by $T_1$ and $T_2$, and these estimates have variances $v_1$ and $v_2$.
The estimates and their variances are what get reported in typical statistical analyses of experiments.

We argue that one useful definition of replication involves the similarity not of estimates $T_i$, but of actual true effects $\theta_i$.
The $T_i$ might differ because of both differences in the underlying true effects, but also due to random chance.
The $\theta_i$ might differ, however, because studies sample participants from different populations, or if there are differences in experimental contexts and conditions between experiments.
Thus, it may make sense to define replication in terms of differences in experiments rather than statistical noise.
Note that there is a connection to the potential outcomes framework in causal inference; for discussion see Wong and Steiner (2017).

In that vein, $\theta_1 = \theta_2$ would be a convenient starting point for defining replication.
However, there is evidence that even in the face of strong theory and sound scientific practice, that studies may still get slightly different results.
Thus, a more practical definition would say studies replicate if $\theta_1 \approx \theta_2$.
For further discussion on this distinction see Hedges and Schauer (2018).

The key distinction here, is that the definition of replication involves the parameters $\theta_i$. 
Analyses of replication involve the estimates $T_i$ and their variances $v_i$, which we use make inferences about how similar the $\theta_i$ are.
This paradigm allows us to evaluate properties of analysis methods.

@@

>>>>

```{r, echo=F, warning=F, message=F}
source("./code/error_rates.R")
```

Programs of research that examine the replicability of research findings have become more common in the social sciences. 
On its face, assessing replication seems trivial: just repeat an experiment and check that you get the same results.
However analyses of these programs reveal that we might mean several different things when we refer to "the same results."
Indeed, the Open Science Collaboration (OSC) (2015) argued, "There is no single standard for evaluating replication success," and then analyzed replicate pairs in a variety of ways.

Despite immediate criticism of the analyses (see Gilbert, Pettigrew, and King, 2015), the core set of OSC analyses has become something of a standard in replication research.
Camerer, et al. (2016) conducted an almost identical research design with the same analytic methods on findings in experimental economics. 
Schweinsberg, et al. (2016), while conducting several---i.e., more than one---replicates of the same findings, aggregated replicate results so that they could use the same analyses as the OSC.
Thus, the idea that replication is impossible to assess has persisted across multiple high-profile replication research efforts.

Given their prevalence in the replication literature, the analyses conducted by the OSC have a few important aspects worth exploring.
The first involves the definitions of replication that they imply. 
For instance, comparing $p$-values versus effect estimates may suggest very different ways we may think about studies' "results" and whether they are "the same."
The second are the properties of the methods used to assess definitions of replication.
For most replication studies, authors seek to make determinations about whether a finding was successfully replicated.
Doing so amid uncertainty---such as in effect estimates---runs into similar issues of null hypothesis tests. 
Thus, unambiguous interpretation of the results of replication research, can depend on the properties of the methods authors use to make these determinations.

In this paper, we explore methods for assessing replication as proposed by the OSC.
We argue that definitions of replication ought to involve effect parameters, rather than $p$-values. 
Using this logic, we then frame the set of analyses common in replication research as null hypothesis tests about replication. 
The properties of these tests are computed under some simplifying assumptions, and shown that there are conditions where they may be suboptimal.

# Model & Framework

One approach to assessing replication proceeds from standard meta-analytic models. 
Meta-analysis represents the results of an experiment by an effect parameter, denoted $\theta$.
This "true" effect reflects what the researcher would observe if they had an infinite sample size.
Effect parameters are typically on a standard scale, such as a standardized mean difference, Fisher-transformed correlation coefficient, or log-odds ratio.

In the context of replication studies, we have at least $k \geq 2$ such findings, so let $\theta_i$ describe the results of the $i$th study.
For programs such as that of the OSC, this invovles $\theta_1$ for the original study and $\theta_2$ for the replicate.
However, we do not actually observe $\theta_1$ and $\theta_2$, and instead we estimate them by $T_1$ and $T_2$, and these estimates have variances $v_1$ and $v_2$.
The estimates and their variances are what get reported in typical statistical analyses of experiments.

We argue that one useful definition of replication involves the similarity not of estimates $T_i$, but of actual true effects $\theta_i$.
The $T_i$ might differ because of both differences in the underlying true effects, but also due to random chance.
The same can be said for the resulting $p$-values---$p_1$ and $p_2$.
The $\theta_i$ might differ, however, because studies sample participants from different populations, or due to differences in contexts and conditions between experiments.
Thus, it may make sense to define replication in terms of differences in experiments rather than statistical noise.
Note that there is a connection to the potential outcomes framework in causal inference; for discussion see Wong and Steiner (2018).

In that vein, $\theta_1 = \theta_2$ would be a convenient starting point for defining replication.
However, there is evidence that even in the face of strong theory and sound scientific practice, that studies may still get slightly different results.
Thus, a more practical definition might posit that studies replicate if $\theta_1 \approx \theta_2$.
For further discussion on this distinction see Hedges and Schauer (2018).
The key distinction here, is that the \textit{definition} of replication involves the parameters $\theta_i$, while \textit{analyses} involve the estimates $T_i$ and their variances $v_i$

@@

# Correspondence in Statistical Significance

One metric used by the OSC to determine if studies successfully replicated to to check that both the original and replicate study corresponded in sign and statistical significance.
That is, findings were deemed to have replcated if both studies found significant effects in the same direction---i.e., both positive---or if both studies found null effects.
This type of metric would seem to answer an important question about replication, particularly of published research which more often involves statistically significant findings.
Indeed, Fisher (1935) noted that "we may say that a phenomenon is experimentally demonstrable when we know how to conduct an experiment which will rarely fail to give us statistically significant results."

However, it has long been noted that differences between significant and nonsignificant findings need not correspond to differences in underlying effect parameters (e.g., Gelman & Stern, 2005). 
Thus, it is entirely possible that $\theta_1 = \theta_2 > 0$, so that the studies replicate exactly, but only $T_1$ is statistically significant.
Alternatively, it could happen that $\theta_1$ and $\theta_2$ are markedly different, but both studies involve statistically significant estimates.

To get a sense of the properties of this test, assume that $\theta_1 = \theta_2 = \theta$, or that the studies replicate exactly.
If this were true, we would want this test to infrequently reject this null hypothesis; in statistical terms, we would want the type I error rate to be low.
The null hypothesis would be rejected if $T_1$ is significant and $T_2$ is either not significant or significant and has an opposite direction (or visa versa). 
Thus, the type I error rate is given by
<!-- \begin{align} -->
<!-- \begin{split} -->
<!-- & P[T_1/\sqrt{v_1} > 1.96]P[T_2/\sqrt{v_2} \leq 1.96] +  -->
<!-- P[T_1/\sqrt{v_1} \leq 1.96]P[T_2/\sqrt{v_2} > 1.96] \\ +  -->
<!-- & P[T_1/\sqrt{v_1} < -1.96]P[T_2/\sqrt{v_2} \geq -1.96] +  -->
<!-- P[T_1/\sqrt{v_1} \geq -1.96]P[T_2/\sqrt{v_2} < -1.96] -->
<!-- \end{split} -->
<!-- \end{align} -->
<!-- Under the model, this can be written as  -->
\begin{align*}
\text{level }
& = [1 - \Phi\left(1.96 - \theta/\sqrt{v_1}\right)] \Phi\left(1.96 - \theta/\sqrt{v_2}\right)\\ 
& + \Phi\left(1.96 - \theta/\sqrt{v_1}\right) [1 - \Phi\left(1.96 - \theta/\sqrt{v_2}\right)]\\ 
& + \Phi\left(-1.96 - \theta/\sqrt{v_1}\right) [1 - \Phi\left(-1.96 - \theta/\sqrt{v_2}\right)]\\
& + [1 - \Phi\left(-1.96 - \theta/\sqrt{v_1}\right)] \Phi\left(-1.96 - \theta/\sqrt{v_2}\right)
\end{align*}

When $\theta_1 = \theta_2 = 0$, so that both studies involve null effects, then the type I error rate is 9.5\%.
However, when $\theta_1 = \theta_2 \neq 0$, the frequency of type I errors depends on the power of each study to detect a non-null effect.
Figure \ref{fig:sig:falsepos} shows the false positive rate of the test as a function of the power of study 1, and the relative size of study 2 to study 1. 
For instance, if study 1 had 60\% power to detect $\theta_1$ and study 2 had 60\% power to detect $\theta_2$---so that the studies are the same size---then the false positive rate will be just under 50\%!

In Figure \ref{fig:sig:falsepos}, we see that the false positive rates are nearly all above the nominal 5\%. 
This is not necessarily an artifact of the scale of the graph. 
Indeed, if study 2 had an infinite sample size, so that $v_2 \to 0$ and $\theta_2/\sqrt{v_2} \to 0$, then the false positive rate of this test for replication converges to one minus the power of the initial study:
\[
\text{Error rate } \to \beta_1
\]
where $1 - \beta_1$ is the power of study 1.
Thus, unless study 1 has at least 95\% power, determining if studies replicated according to correspondence of sign and statisitical significance can result in type I error rates well above 5\%.

```{r, echo=F, message=F, fig.cap = "False Positive Rate of Test for Replication. This plot displays the false positive rate of determining replication according to correspondence of sign and statistical significance as a function of the size of study 1 and study 2. Each colored line corresponds to study 2 being a multiple of the size of study 1. The dashed line corresponds to a 5\\% false positive rate. \\label{fig:sig:falsepos}"}
falsepos_plt
```

Conversely, if we assume $\theta_1 \neq \theta_2$, then we would want the test to reject the null hypothesis frequently---that is, we would want a low type II error rate.
